Numerical simulation of the Cordilleran ice sheet through the last glacial cycle

### 2014-06-11 cordillera-cycle.tex
Deglaciation history first shot.

Deglaciation happens first in the southern half and then in the northern half
of the ice-covered area (Fig.~\ref{fig:deglac}). Retreat patterns are different
than reconstructed over central BC, however they match Martin's reconstruction
of Liard lowland "unzipping" pretty well. In the GRIP-driven simulation, the
patterns of de-glaciation and Younger Dryas re-advance match with recent
cosmogenic exposure dates.

>>>>

In the North American Cordillera alike other glaciated regions, the large
majority of the glacial geologic record relates to the last few millennia of
glacier cover, most of the older evidence having been overridden by ice retreat
retreat during the deglaciation.

In our simulations, the timing of the last glacial maximum, and in turn that of
the deglaciation, depends on the palaeo-temperature time-series used to force
the model (Figs.~\ref{fig:timeseries} and ~\ref{fig:deglacseries}). The
6\,\unit{km}-resolution simulation forced by the EPICA record reaches a maximal
ice volume at 17.3\,\unit{kyr}, followed by a continuous deglaciation between
about 17 and 11\,\unit{kyr}. That driven by the GRIP record reaches maximal
volume at 19.1\,\unit{kyr}, and produce a deglaciation in three steps,
including a first stand-still between 16.6 and 14.5\,\unit{kyr}, and a late
glacial re-advance between 12.6 and 11.6\,\unit{kyr}.

Although the timing differs, the patterns of glacier retreat are relatively
consistent between the two simulations (Figs.~\ref{fig:deglac}). In both
simulations, the southern half of the modelling domain, including the Puget
Lowland, the Coast and Rocky Mountains and the Interior Plateau of British
Columbia, are completely deglaciated at a time where there remain significant
ice cover on the Skeena, Selwyn, MacKenzie, Wrangell and Saint-Elias Mountains
in the northern half of the domain. At this stage, a late glacial re-advance
occurs in the GRIP simulation. Both simulations then show an unzipping of the
Liard Lowland, followed by ice retreat towards the Selwyn Mountains and the
Skeena Mountains, which host the last remain of the palaeo-ice sheet.

On the Interior Plateau of British Columbia, modelled deglaciation is rapid
and northwards, resulting in the presence of a slow-flowing ice lobe on the
plateau, while the surrounding mountain ranges are already deglaciated
(Fig.~\ref{fig:profiles}). Although this supports the top-down (?)
deglaciation hypothesis for central British Columbia, modelled, southwards,
last flow directions do not agree with the local lineation patterns
(Fig.~\ref{fig:lastflow}). This could indicate that the lineations are older,
or that the modelled last glacial maximum ice sheet is too extensive to the
east, resulting in modelled deglaciation flow patterns different from those
observed in the geomorphology.
